---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# narwbio

<!-- badges: start -->
<!-- badges: end -->

The package uses the bioenergetic model implemented in:
Lehoux, C., Plourde S., and Lesage, V. 2020. Significance of dominant zooplankton species to the North Atlantic Right Whale potential foraging habitats in the Gulf of St. Lawrence: a bio-energetic approach. DFO Can. Sci. Advis. Sec. Res. Doc. 2020/033. iv + 44 p.

The original model was published by:
Gavrilchuk, Katherine, et al. "Foraging habitat of North Atlantic right whales has declined in the Gulf of St. Lawrence, Canada, and may be insufficient for successful reproduction." Endangered Species Research 44 (2021): 113-136.

AND 

Gavrilchuk, K., Lesage, V., Fortune, S., Trites, A.W., and Plourde, S. 2020. A mechanistic
approach to predicting suitable foraging habitat for reproductively mature North Atlantic right
whales in the Gulf of St. Lawrence. DFO Can. Sci. Advis. Sec. Res. Doc. 2020/034. iv + 47
p.

The Lehoux et al. implementation of this model include more species in the total biomass and the ENet definition is different. The NARW is expected to target one depth layer where its Enet is maximum. Enet is calculated at each position. 

This package works with the vertical distribution of C.hyp that includes its development index (Comtois et al. in prep). This vertical distribution tends to decrease the level of overestimation which was a problem with Lehoux et al. 2020. We are discussing a publication for the valdiation of our vertical distribution models for 2023.

If you have Calanus abundances instead of dry weight, use Plourde et al. 2019 for conversion. Can be integrated in the future...

The models from Lehoux et al. are priorize when the Region is set to gsl and season and Day_Night are available, otherwise, models from Plourde et al. 2019 are used. 

The package now accomodate and export for different state resting, pregnant and lactating.
and sensitivity of parameters is now allowed

New column is exported for minimum density of preys.


## Installation

You can install the development version of narwbio like so:
since this is not a public repository, your R studio must be connected to Github.
First use:
```{r dred}
credentials::set_github_pat() 
```


if you have already entered a token on your computer, this should not do anything visible and then install the package: 
```{r instal}
devtools::install_github("clehoux/narwbio")
library(tidyverse)

```



## Example

```{r exem}
dat <-  data.frame(Label="label1", Z_Cf.Cg4.6_mgm2=1000, Z_Ch4.6_mgm2=10000,Z_Pcal.Micro.Scol_mgm2=100,Z_Tem.Eury_mgm2=100, Day_Night="D", Zstation=100, season="esum", MeshSize=333)

library(narwbio)
res_out<- execute_bioenergy(dat, #dataframe like zooimage
                            cfin = "Z_Cf.Cg4.6_mgm2", #column names for cfin dw
                            chyp = "Z_Ch4.6_mgm2", # column names for chyp dw
                            psca = "Z_Pcal.Micro.Scol_mgm2", 
                            temo = "Z_Tem.Eury_mgm2",
                            cfinZ = NULL, #optionnaly in cases of 3D model output mostly for Nicolas Lecorre
                            chypZ = NULL,
                            pscaZ = NULL,
                            temoZ = NULL,
                            units = "mg", #units of dry weight
                            ID = "Label",# unique Idientifier. for the whole zooimage dataset use Label and SampleAnalysis pasted together.
                            change_150m = T) # T for th Gavrilchuk model , F for exploration.

res_out

```


## Example to get minimum density

```{r exem2}
dat <-  data.frame(Label="label1", Z_Cf.Cg4.6_mgm2=1000, Z_Ch4.6_mgm2=10000,Z_Pcal.Micro.Scol_mgm2=100,Z_Tem.Eury_mgm2=100, Day_Night="D", Zstation=100, season="esum", MeshSize=333)

library(narwbio)
res_out<- execute_bioenergy(dat, #dataframe like zooimage
                            cfin = "Z_Cf.Cg4.6_mgm2", #column names for cfin dw
                            chyp = "Z_Ch4.6_mgm2", # column names for chyp dw
                            psca = "Z_Pcal.Micro.Scol_mgm2", 
                            temo = "Z_Tem.Eury_mgm2",
                            cfinZ = NULL, #optionnaly in cases of 3D model output mostly for Nicolas Lecorre
                            chypZ = NULL,
                            pscaZ = NULL,
                            temoZ = NULL,
                            units = "mg", #units of dry weight
                            ID = "Label",# unique Idientifier. for the whole zooimage dataset use Label and SampleAnalysis pasted together.
                            change_150m = T,
                            state="pregnant",
                            param="mean") # T for th Gavrilchuk model , F for exploration.

resa<- expand_grid(res_out, Zsampled=seq(10,500, 10)) %>%  filter(Zsampled <=Zstation) %>%  mutate(percZ_stn=Zsampled/Zstation*100) 

resa$prop_biom_cum= predict(mcfin_glac_c_gsl_esum_D, resa, type="response")
test<-  resa %>% ungroup() %>% filter(Zstation >10) %>%  arrange(Label, Zsampled) %>% group_by(Label) %>%  mutate(prop_biom=c(first(prop_biom_cum),diff(prop_biom_cum, lag = 1)))


maxdf <- test %>% filter(ED.max_Z==Zsampled) %>%  mutate(MinDensity_cfinc5_gm3=Minimum_prey_density_gm3,
                                                         MinDensity_cfinc5_mgm3=MinDensity_cfinc5_gm3*1000,
                                                         Min_cfinc5_indm3=MinDensity_cfinc5_mgm3/0.3,
                                                         Min_cfinc5_indm2=Min_cfinc5_indm3/(prop_biom/10))#porportion is calculated on 10m3

library(ggthemes)
ggplot(maxdf %>%  filter(Zstation < 400), aes(x=Zstation, y=Min_cfinc5_indm2/1000)) +geom_point() +geom_hline(yintercept=10)+theme_few() + ylab("Min Cfin C5 10³ ind/m²") +scale_y_continuous(trans="log10")


```


